UZH IT \& AI Emissions Analysis

### 2024-11-28 doc.tex
Add information about phones, other devices, networking, cloud footprint.

\subsection*{UZH parameters}

We have access to a spreadsheet (\texttt{2024-11-13\_Berechnung\_THG\_IKT\_2024.xls}) with
counts and emissions factors for various types of IT equipment at UZH. It cites approximately
10,000 laptops and monitors, 1,000 servers, 2,400 mobile phones,  150 classroom PCs, and a few
hundred printers, copiers, and projectors.

By contrast, Wikipedia \parencite{wikiuzh} cites approximately 28,000 students and 10,000 staff
at UZH. Evidently the spreadsheet only lists university assets, not personal devices.

@@

\subsection*{Embodied emissions}
The estimation of embodied emissions introduces additional assumptions and uncertainties into
the model. The secretiveness of major cloud providers makes it especially difficult
to estimate the embodied emissions of cloud infrastructure. Our initial proposal was to ignore
embodied emissions altogether, but that leaves a big gap.

@@

\begin{table}[h]
\centering
\begin{tabular}{|r|l|}
\hline
Production \& Disposal & \\
(kg CO$_2$e) & Source \\ \hline
121 & Ecoinvent  \\ \hline
104 & \textcite{teehan2013} (low) \\ \hline
338 & \textcite{teehan2013} (high) \\ \hline
244 & \textcite{rarecoil} \\ \hline
202 & \textcite{unctadder2024} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:laptops}
\end{table}

The Ecoinvent numbers appear to be outliers on the low end.

The most comprehensive data comes from \textcite{rarecoil}, which lists manufacturer
information for over 90 popular devices. Mean supply chain emissions are
244 kg CO$_2$e with a standard deviation of 128 kg. We model laptop emissions using a normal
distribution with those characteristics.

@@

\begin{table}[h]
\centering
\begin{tabular}{|r|l|}
\hline
Production \& Disposal & \\
(kg CO$_2$e) & Source \\ \hline
238 & Ecoinvent \\ \hline
303 & \textcite{teehan2013} \\ \hline
403 & \textcite{unctadder2024} \\ \hline
289 & \textcite{dellpcf} \\ \hline
277 & \textcite{boavizta} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:desktops}
\end{table}

Most estimates are in the range of 250 - 400 kg CO$_2$e. The most detailed data comes from
Dell's Product Carbon Footprints (\textcite{dellpcf}), which they publish for each of their products.
A sample of 15 desktops has mean supply chain emissions of 289 kg CO$_2$e with $\sigma = 80$ kg.
We model desktop emissions using a normal distribution with those characteristics.

@@

\begin{table}[h]
\centering
\begin{tabular}{|r|l|}
\hline
Production \& Disposal & \\
(kg CO$_2$e) & Source \\ \hline
383 & \textcite{teehan2013} \\ \hline
1252 & \textcite{davy2021} (Dell) \\ \hline
1912 & \textcite{davy2021} (EC2) \\ \hline
899 & \textcite{boavizta} (medium server)\\ \hline
\end{tabular}
\label{tab:embodied_emissions:servers}
\end{table}

\textcite{davy2021} reports data for many Dell servers and estimates for different types of EC2 instances.
The former have mean supply chain emissions of 1252 kg CO$_2$e with $\sigma = 330$ kg, the latter
1912 kg with $\sigma = 885$ kg. We model server emissions using the distribution corresponding to Dell
servers.

@@

@@

>>>>

\subsection*{UZH population}

We have access to a spreadsheet (\texttt{2024-11-13\_Berechnung\_THG\_IKT\_2024.xls}) with
counts and emissions factors for various types of IT equipment at UZH. It reports approximately
10,000 laptops and monitors, 1,000 servers, 2,400 mobile phones,  150 classroom PCs, and a few
hundred printers, copiers, and projectors.

By contrast, \textcite{wikiuzh} reports approximately 28,000 students and 10,000 staff
at UZH. Evidently the spreadsheet only lists university assets, not personal devices.

@@

\subsection*{Embodied emissions}
The estimation of embodied emissions introduces additional assumptions and uncertainties into
the model. Our initial project proposal ignored embodied emissions altogether, but that leaves
too big a gap.

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
Production \& Disposal & \\
(kg CO$_2$e) & Source \\ \hline
121 & Ecoinvent  \\ \hline
104 & \textcite{teehan2013} (low) \\ \hline
338 & \textcite{teehan2013} (high) \\ \hline
244 & \textcite{rarecoil} \\ \hline
202 & \textcite{unctadder2024} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:laptops}
\end{center}

The Ecoinvent numbers appear to be outliers on the low end.
The most comprehensive data comes from \textcite{rarecoil}, which lists manufacturer
information for over 90 popular devices. Mean supply chain emissions are
244 kg CO$_2$e with a standard deviation of 128 kg.\
\assumption{We model laptop supply chain emissions using a normal distribution with those characteristics.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
Production \& Disposal & \\
(kg CO$_2$e) & Source \\ \hline
238 & Ecoinvent \\ \hline
303 & \textcite{teehan2013} \\ \hline
403 & \textcite{unctadder2024} \\ \hline
289 & \textcite{dellpcf} \\ \hline
277 & \textcite{boavizta} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:desktops}
\end{center}

Most estimates are in the range of 250 - 400 kg CO$_2$e. The most detailed data comes from
Dell's Product Carbon Footprints (\textcite{dellpcf}), which they publish for each of their products.
A sample of 15 desktops has mean supply chain emissions of 289 kg CO$_2$e with $\sigma = 80$ kg.
\assumption{We model desktop supply chain emissions using a normal distribution with those characteristics.}

@@

\begin{center}
\begin{tabular}{|r|l|}
\hline
Production \& Disposal & \\
(kg CO$_2$e) & Source \\ \hline
383 & \textcite{teehan2013} \\ \hline
1252 & \textcite{davy2021} (Dell) \\ \hline
1912 & \textcite{davy2021} (EC2) \\ \hline
899 & \textcite{boavizta} (medium server)\\ \hline
\end{tabular}
\label{tab:embodied_emissions:servers}
\end{center}

\textcite{davy2021} reports data for many Dell servers and estimates for different types of EC2 instances.
The former have mean supply chain emissions of 1252 kg CO$_2$e with $\sigma = 330$ kg, the latter
1912 kg with $\sigma = 885$ kg.
\assumption{We model server supply chain emissions using the distribution corresponding to Dell servers.}

\subsubsection*{Smartphones}

\begin{center}
\begin{tabular}{|r|l|}
\hline
Production \& Disposal & \\
(kg CO$_2$e) & Source \\ \hline
68 & \textcite{googlepixel8} (Pixel 8)\\ \hline
54 & \textcite{appleiphone13} (iPhone 13)\\ \hline
50 & \textcite{unctadder2024} \\ \hline
50 & \textcite{lovehagen2023} \\ \hline
\end{tabular}
\label{tab:embodied_emissions:phones}
\end{center}

\assumption{We model phones as having mean supply chain emissions of 50 kg CO$_2$e with $\sigma = 10$ kg.}

\subsubsection*{Other devices}

We have found less data for other types devices. For now we model them with averages based
on a handful of sources.

\begin{description}
    \item[Computer monitors] 344 kg CO$_2$e (\textcite{teehan2013}, \textcite{dellpcf})
    \item[Conference room displays] 753 kg CO$_2$e (scaling monitor by $(40''/27'')^2$)
    \item[Printer/copier stations] 1167 kg CO$_2$e (Ecoinvent)
    \item[Network equipment] We did not find reliable lifecycle assessments for routers and switches,
      and have no information about UZH network architecture. The vast majority ($80-95\%$) of
      lifecycle emissions of network gear stem from usage (\textcite{cisco2024}, \textcite{jacob2023}).
      For now, we model network equipment as a 5\% overhead on servers (1 router or switch for 20 servers).
      (A typical data center rack contains 42 1U servers and one switch.)
\end{description}

@@

The secretiveness of major cloud providers makes it difficult to estimate the embodied emissions of cloud
infrastructure. The best resource we have been able to find is the Datavizta API (\textcite{boavizta}).

Lacking information about UZH's workloads, we model a hypothetical UZH cloud footprint as follows:
\begin{itemize}
    \item One-for-one replacement of one campus server with one Azure D8S\_v3 instance (2x Intel
      Xeon, 8 cores, 32 GB RAM, 2TB SSD) (\cite{msftvms}).
      \item Deployment to Microsoft Azure's ``Switzerland North'' region .
      \item Estimation of annual emissions via Datavizta, assuming constant 50\% load: 31 kg CO$_2$e for usage,
      97 kg CO$_2$e for manufacturing.
\end{itemize}

@@

\subsection*{Other IT components}

For now, we do not attempt to model changes in network infrastructure or utilization due to
move to the cloud or increased LLM use. See \textcite{jacob2023} for an overview of
network sustainability.

\subsection*{Potential AI benefits}
